---
title: "aspen-fire_frp"
output: html_document
date: "2023-05-01"
---

## Aspen Controls on Satellite-derived Fire Radiative Power (FRP)

Setup the data frame and workspace:

```{r setup}

knitr::opts_chunk$set(echo = TRUE)

# LIBRARIES AND GLOBAL VARS
library(tidyverse)
library(sf)
library(ggpubr)
library(classInt)
library(ggcorrplot)
library(rstatix)
library(gt)

getwd()

proj <- st_crs("ESRI:102039")

# DATA

# Fire Radiative Power (FRP) observations (plots) w/ attributes from GEE
frp <- st_read('../../data/spatial/mod/vnp14img_west_spatial_w_attr.gpkg') %>%
 # tidy the frame
 mutate(daynight = as.factor(daynight),
        acq_year = lubridate::year(acq_date),
        aspen = if_else(aspen_pct > 0, 1, 0)) %>%
 # Filter any missing FRP data and only for "nominal" or "high" confidence observations
 filter(frp > 0,
        confidence != "l") %>%
 # Convert back to centroid (points) and ensure correct project
 st_centroid() %>% 
 st_transform(proj)

gc() # garbage clean

```

```{r}
glimpse(frp)
```

Grab a dataframe with forested observations and with quantile breaks:

```{r}
brks <- classIntervals(frp$aspen_pct, style='headtails')
brks <- brks$brks

df <- frp %>% 
 mutate(aspen_qnt = factor(cut(aspen_pct, breaks=brks)),
        aspen_qnt = if_else(is.na(aspen_qnt), factor(0), aspen_qnt),
        mtbs_id = factor(mtbs_id)) %>%
 filter(lcms_mode == 1) # forested plots (1=Trees, 3=Shrub/Tree Mix, 4=Grass/Trees Mix, 5=Barren/Trees Mix)
```

# Data exploration and summary statistics

Distribution of FRP Observations:

```{r message=F, warning=F, fig.height=2.5, fig.width=6}

p1 <- ggplot(data=frp, aes(x=frp)) +
  geom_histogram(bins=30)  +
  labs(y="Count",x="frp") +
  theme_minimal(5) +
  theme(axis.text = element_text(size=7),
        axis.text.x = element_text(size=5),
        axis.text.y = element_text(size=5))
p2 <- p1 +
  scale_x_continuous(trans="sqrt") +
  labs(x="sqrt(frp)",y="")
p3 <- p1 +
  scale_x_continuous(trans="log10") +
  labs(x="log(frp)",y="")

(arr <- ggarrange(p1,p2,p3,nrow=1,ncol=3))

# ggsave(arr,file="../../figures/frp/west_frp_distributions.png",dpi=150)
rm(p1,p2,p3,arr)

```

Plot the day/night observations on a three month interval.

```{r message=F}

p1 <- ggplot(data=frp, aes(x=cut(acq_date, breaks="3 month"), y=frp, fill=factor(daynight))) +
  geom_boxplot(outlier.size = 0.8) +
  scale_y_continuous(trans="log10") +
  scale_fill_manual(values=c("#ef8a62","#999999"), labels=c("Daytime Obs.","Nighttime Obs.")) +
  labs(x="Acquisition Date", y="frp (gW)", fill="") +
  theme_bw(10) +
  theme(axis.text.x = element_text(angle=35,hjust = 1),
        legend.position = "top")
p1

rm(p1)

```

Plotting aspen / no-aspen and FRP across time:

```{r message=F}

# Plots

p1 <- ggplot(data=df, aes(x=cut(acq_date, breaks="3 month"), y=frp, fill=factor(aspen))) +
  geom_boxplot(outlier.size = 0.6) +
  scale_y_continuous(trans="log10") +
  scale_fill_manual(values=c("#999999","#238443"), labels=c("No Aspen","Aspen")) +
  labs(x="Acquisition Date", y="frp (gW)", fill="Aspen Presence:") +
  facet_wrap(~daynight, nrow=2) +
  theme_bw(10) +
  theme(axis.text.x = element_text(angle=35,hjust = 1),
        legend.position = "top")
p1

# ggsave(p,file="../../figures/frp/frp_AspenNoAspen_boxplot.png",dpi=300)
 
rm(p1)

```

Produce the same plots but using the quantiles:

```{r message=F}

cols <- c('#ffffcc','#c2e699','#78c679','#238443','#005824')

p1 <- ggplot(data=df, aes(x=cut(acq_date, breaks="3 month"), y=frp, fill=aspen_qnt)) +
  geom_boxplot(outlier.size = 0.6) +
  scale_y_continuous(trans="log10") +
  scale_fill_manual(values=cols, 
                    labels=c("None","(0-6.96)","(6.96-28.24)","(28.24-54.69)","(54.69-99.24)")) +
  labs(x="Acquisition Date", y="frp (gW)", fill="Aspen Component (%):") +
  facet_wrap(~daynight, nrow=2) +
  theme_bw(10) +
  theme(axis.text.x = element_text(angle=35,hjust = 1),
        legend.position = "top")
p1

# ggsave(p1,file="../../figures/frp/FRP-Aspen_Quantile_byWeek.png",dpi=300)

rm(p1)
```

Test the relationship with a GAM.


```{r}

library(mgcv,quietly=T)
library(tidymv,quietly=T)
ctrl <- mgcv::gam.control(nthreads = 6) # use 6 parallel threads, reduce if fewer physical CPU cores

df_ <- df %>% filter(pct_tree > 50)

# Run a GAM
gam.fit <- 
  mgcv::gam(log(frp) ~ s(aspen_pct, k=15) + s(vpd_max) + s(slope),
  method="REML",
  data=df_, 
  control=ctrl
)

summary(gam.fit)

# Plot the smooths
plot_smooths(gam.fit, aspen_pct) +
 theme(legend.position = "right") +
 labs(x="Aspen %") +
 theme_bw(14)

plot_smooths(gam.fit, vpd_max) +
 theme(legend.position = "right") +
 labs(x="Max VPD") +
 theme_bw(14)

plot(gam.fit)

```

Statistical tests for significance between groups / aspen percent more broadly:

```{r message=F, warning=F}
# Summary stats
table <- df %>%
 select(aspen_qnt, frp) %>%
 st_set_geometry(NULL) %>% 
 as_tibble() %>%
 group_by(aspen_qnt) %>%
 get_summary_stats(frp, type = "mean_sd") %>%
 rename(Aspen_Pct = aspen_qnt,
        N = n,
        Mean = mean,
        StDev = sd) %>%
 select(Aspen_Pct, N, Mean, StDev) %>%
 gt() %>%
 tab_header(title = "FRP ~ Aspen %") %>%
 tab_style(
  style = list(cell_fill(color = "#b2f7ef"),
  cell_text(weight = "bold")),
  locations = cells_body(columns = Mean))
table

# gtsave(table, "../../figures/frp/Table_FRP_AspenPct_Summary.png")

rm(table)
```

```{r warning=F, message=F}
# ANOVA test of significance
aspen.anova <- anova(lm(log(frp) ~ aspen_qnt, data = df))
aspen.anova

# Run a GAM to get an estimate of the effects for each group
glm.fit <- glm(log(frp) ~ aspen_qnt, data = df)
summary(glm.fit)
sjPlot::tab_model(glm.fit)

# Create the effects plot
p <- sjPlot::plot_model(glm.fit, show.values = TRUE, value.offset = .3) + 
 labs(y="Model Estimate",x="Aspen Component",title="FRP ~ Aspen Component (%)") +
 scale_x_discrete(labels=c("Aspen_qt(0,6.87]" = "0-6.9%", "Aspen_qt(6.87,28]" = "6.9-28%", 
                           "Aspen_qt(28,54.6]" = "28-54.6%", "Aspen_qt(54.6,99.2]" = "54.6-99.2%")) +
 theme_bw(12) +
 theme(axis.text.x = element_blank(),
       axis.ticks.x = element_blank(),
       axis.text.y = element_text(angle=30,hjust = 1))
p
# ggsave(p,file="../../figures/frp/FRP-Aspen_Quantile_GLM_estimates.png",dpi=120,height=5,width=4)
rm(p,glm.fit,aspen.anov)
```

```{r message=F}

p1 <- ggplot(data=df%>%filter(confidence=='h'), aes(x=aspen_qnt, y=frp, fill=daynight)) +
 geom_boxplot(outlier.size = 0.8) +
 scale_y_continuous(trans="log10") +
 scale_fill_manual(values=c("#ef8a62","#999999"), labels=c("Daytime Obs.","Nighttime Obs.")) +
 labs(y="log(FRP)",x="Aspen Component") +
 theme_bw(10) +
 theme(legend.position = "top")
p1

# ggsave(p1,file="../../figures/frp/FRP_Aspen_Breaks_boxplot.png",dpi=150)

rm(p1)

```

Plot of FRP across major cover types:

```{r message=F}
# LANDFIRE table
codes <- read_csv("../../../../data/landcover/LANDFIRE/LF2016_EVT_200_CONUS/CSV_Data/LF16_EVT_200.csv") %>%
 mutate(lfevt = factor(VALUE)) %>%
 select(lfevt,EVT_PHYS,EVT_GP_N)
# Majority type
lfevt <- read_csv("../../data/tabular/mod/LFEVT16/vnp14img_lf_evt_majority.csv") %>%
 mutate(lfevt = factor(lfevt_majority)) %>%
 select(uid,lfevt) %>%
 left_join(codes,by="lfevt")

# Join back to FRP data
df <- df %>% left_join(lfevt, by="uid")

summary(lfevt$lfevt)
glimpse(df)

rm(codes,lfevt)
```


```{r, message=F}
# Filter to LFEVT codes with > 100 observations
df.c <- df %>%
  na.omit() %>%
  group_by(lfevt) %>%
  filter(n() > 100)

# Plot groups
ggplot(data=df.c, aes(x=EVT_GP_N, y=FRP)) +
 geom_boxplot() +
 scale_y_continuous(trans="log10") +
  theme_bw(10) +
  theme(axis.text.x = element_text(angle=35,hjust = 1),
        legend.position = "top")
```

```{r}
df.c <- df %>%
 st_set_geometry(NULL) %>% as_tibble() %>%
 mutate(
  Dominant = if_else(Conifer == 100 & Aspen < 25, "Conifer", NA),
  Dominant = if_else(Aspen > 75, "Aspen", Dominant),
  Dominant = factor(Dominant)
 )
head(df.c,10)
summary(df.c$Dominant)
# Plot groups
ggplot(data=df.c %>% filter(Dominant=="Aspen"|Dominant=="Conifer"), aes(x=Dominant, y=FRP)) +
 geom_boxplot() +
 scale_y_continuous(trans="log10") +
  theme_bw(10) +
  theme(axis.text.x = element_text(angle=35,hjust = 1),
        legend.position = "top")
```

Statistical test:

```{r}
temp <- df.c %>% filter(Dominant=="Aspen"|Dominant=="Conifer")

# Anova test
aspen.anov <- anova(lm(log(FRP) ~ Dominant, data = temp))
aspen.anov

# Run a GLM to get an estimate of the effects for each group
glm.fit <- glm(log(FRP) ~ Dominant, data = temp)
summary(glm.fit)
exp(glm.fit$coefficients)
sjPlot::tab_model(glm.fit)
rm(temp)
```

GAM of Aspen percent:

```{r}
library(mgcv,quietly=T)
library(tidymv,quietly=T)
ctrl <- mgcv::gam.control(nthreads = 6) # use 6 parallel threads, reduce if fewer physical CPU cores
df <- df %>% mutate(Event_ID = factor(Event_ID))
# Run a GAM
gamfit1 <- 
  mgcv::gam(log(FRP) ~ s(Aspen, k=15),
  method="REML",
  gamma=1.2,
  data=df, 
  control=ctrl
)
###Check the model summaries (including k.check) and anova
summary(gamfit1) #Standard model summary
print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~GAM-Check~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
gam.check(gamfit1) #Check the basis dimension choice for smooths (k)
print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ANOVA~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
anova.gam(gamfit1) #anova for approximate significant of smooth terms
# Plot the smooths
plot_smooths(gamfit1, Aspen) +
 theme(legend.position = "top") +
 labs(x="Aspen %") +
 theme_bw(14)
```

